### 2020-03-05 Common/chapters/conclusions.tex
Johan and Julien: Thesis layout updated until transfer learning :)

\section{Conclusion}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.\\
The starting point was the reproduction of Song et al.'s experiment~\cite{07}. Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods were working well. This resulted in a solid baseline that was exploited in order to take part in the "SPIE-AAPM-NCI Prostate MR Classification Challenge”, also called PROSTATEx challenge. Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge. This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge, which confirms the robustness of the latter.  \\
Then, the work focused on overcoming one of the main issues in deep learning: the lack of data. To achieve this, transfer learning, as well as more common techniques such as data augmentation, were applied. The particularity of our transfer learning implementation is the intermediate frozen steps. During the latter, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does. Experiments showed that: XXXXXX RESULTS XXXXXX.

Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), an all-in-one training and testing file which can be adapted to new models and datasets, an end-to-end transfer learning pipeline. All these elements can serve as a baseline or as additions to future projects.

xxxxx

>>>>

\section{Conclusion}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.\\
The starting point was the reproduction of Song et al.'s experiment~\cite{07}. Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods were working well. This resulted in a solid baseline that was exploited in order to take part in the "SPIE-AAPM-NCI Prostate MR Classification Challenge”, also called PROSTATEx challenge. Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge. This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge, which confirms the robustness of the latter.  \\
Then, the work focused on overcoming one of the main issues in deep learning: the lack of data. To achieve this, transfer learning, as well as more common techniques such as data augmentation, were applied. The particularity of our transfer learning implementation is the intermediate frozen steps. During the latter, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does. Experiments showed that: XXXXXX RESULTS XXXXXX.\\
Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), an all-in-one training and testing file which can be adapted to new models and datasets, an end-to-end transfer learning pipeline. All these elements can serve as a baseline or as additions to future projects.\\
xxxxx

### 2020-03-05 Common/chapters/deep_learning.tex
Johan and Julien: Validation process corrected + nifti reference added
Johan and Julien: Thesis layout updated until transfer learning :)

\subsection{Perceptrons}
\label{perceptron}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Perceptrons are the main components of neural networks. They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}. Today, they are called "artificial neurons" or simply "neurons".\\
The output $a$ of a perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
\begin{equation}
\label{perceptron_equation}
a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{j})
\end{equation}

Schematically speaking, a perceptron can be represented as on figure \ref{perceptron_model}. Each input is multiplied with its corresponding weight. The sum of this result then goes through a non-linear function, called "activation function". This activation function acts like a threshold that determines the proportion of the result that goes further in the network. There exist multiple activation functions (see \ref{activation_functions}). It is extremely important to use non-linear functions instead of a linear functions. In fact, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}). Consequently, if only linear functions are used throughout the network, linear outputs are given as inputs to other linear functions. Since the composition of two linear functions is itself a linear function, assembling perceptrons to create neural networks of multiple layers would not make sense anymore.

@@

\subsection{Activation functions}
\label{activation_functions}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer. Activation functions must be non-linear in order to approximate extremely complex functions. In fact, neural networks are considered as universal approximators. Hornik et al. claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense" \cite{17}. According to Thomas Epelbaum \cite{18}, the most commonly used activation functions are:\\\\
 \noindent \textbf{Sigmoid function}\\
 The sigmoid function is defined as:
 \begin{equation}
  g(x) = \frac{1}{1+e^{-x}}
 \end{equation}
 Its derivative is:
 \begin{equation}
 g'(x) = g(x)(1-g(x))
 \end{equation}

@@

There exist multiple cost functions. For a regression problem, a commonly used loss function is the mean square error:
\begin{equation}
J(\hat{y}, y) = \frac{1}{m}[\sum_{i=1}^{m} (\hat{y}^{(i)} - y^{(i)})^{2}]
\end{equation}
For classification problems, the cross entropy function is regularly used. We distinguish the binary classification where the number of classes $n_{y}$ = 2 from the multiclass classification where $n_{y}$ > 2. In the case of binary classification, the cross entropy is:
\begin{equation}
J(\hat{y}, y) = -\frac{1}{m}\sum_{i=1}^{m} [y_{i}*log(\hat{y}_{i}) + (1-y_{i})*log(1-\hat{y_i})]
\end{equation}
In the case of multiclass classification, the categorical crossentropy is:
\begin{equation}
J(\hat{y}, y) = - \sum_{i=1}^{n_{y}} \sum_{j=1}^{m} (y_{ij}*log(\hat{y}_{ij}))
\end{equation}

Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weights in order to approach the minimum of the function. Therefore, deep learning problems can be considered as optimization problems. Solutions to these problems can be found using the gradient descent algorithm during backpropagation.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$. Generally speaking, "the intuition behind the backpropagation algorithm is as follows. Given a training example $(x^{(i)}, y^{(i)})$, we will first run a forward pass to compute all the activations throughout the network, including the output value of the network. Then, for each node $i$ in layer $l$, we would like to compute an "error term" $\partial^{(l)}_{i}$ that measures how much that node was "responsible" for any errors in our output. For an output node, we can directly measure the difference between the network’s activation and the true target value, and use that to define $\partial^{(n_{l})}_{i}$(where layer $n_{l}$ is the output layer). How about hidden units? For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.

In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting parameters of the model. The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network. The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ at point $x$ is given by:
\begin{equation}
\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{1}}]
\end{equation}
The gradient shows how much all parameters that constitute x need to change to minimize the function. In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$. The computation of all these gradients relies on the "chain rule". In the case of weights, the chain rule is:
\begin{equation}
\frac{\partial J}{\partial w_{jk}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial w_{jk}^{l}}
\end{equation}
Similarly, the chain rules has to be applied to the biases:
\begin{equation}
\frac{\partial J}{\partial b_{j}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial b_{j}^{l}}
\end{equation}
Once the gradients of each parameter are computed, these parameters are updated. The weights update is described by the following equation:
\begin{equation}
W^{[l]} = W^{[l]} - \alpha * \frac{\partial J}{\partial W^{[l]}}
\end{equation}
The biases update corresponds to:
\begin{equation}
b^{[l]} = b^{[l]} - \alpha * \frac{\partial J}{\partial b^{[l]}}
\end{equation}
%The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be tuned manually.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Convolutional Neural Networks (CNN) are a specific type of deep neural networks. They are particular in that they contain layers which perform a mathematical operation named convolution on the input data. CNNs are mostly used in image and video analysis.

To perform a convolution, numerical input data and a "filter" are required. A filter can be seen as a $f$x$f$ numerical patch that moves across the entire input. It first moves horizontally until reaching the right-most border of the image. Then, it goes down a cell and starts from the border on the left-hand side. This process is repeated until the filter reaches the bottom-right corner of the image, which marks the end of the convolution. At each step, the dot product between the filter and the part of the input covered by the filter is computed. Figure \ref{fig:convolution} showcases a simple convolution which aims at finding vertical lines in a black and white image. For example, the output of the first step of the convolution (in red) is computed by evaluating the dot product between the blue filter and the red part of the input image: 
\begin{equation}
\begin{gathered}
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}0
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}= -2
\end{gathered}
\end{equation}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/convolution.png}
\caption{Convolutions - Basic convolution in a CNN}
\label{fig:convolution}
\end{figure}

Different parameters can change the way a convolution behaves. First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom. In this case, the so-called "stride" is equal to 1. Other applications could rely on a bigger stride. Furthermore, the previous example reduced the output size of $3$x$3$ in proportion to the initial input size of $5$x$5$. To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s. Three ways of padding images are commonly used as shown on figure \ref{fig:convolution_padding}:
\begin{itemize}
	\item \textbf{Valid:} The input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.

>>>>

\subsection{Perceptrons}
\label{perceptron}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Perceptrons are the main components of neural networks. They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}. Today, they are called "artificial neurons" or simply "neurons".\\
The output $a$ of a perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
\begin{equation}
\label{perceptron_equation}
a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{j})
\end{equation}
Schematically speaking, a perceptron can be represented as on figure \ref{perceptron_model}. Each input is multiplied with its corresponding weight. The sum of this result then goes through a non-linear function, called "activation function". This activation function acts like a threshold that determines the proportion of the result that goes further in the network. There exist multiple activation functions (see \ref{activation_functions}). It is extremely important to use non-linear functions instead of a linear functions. In fact, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}). Consequently, if only linear functions are used throughout the network, linear outputs are given as inputs to other linear functions. Since the composition of two linear functions is itself a linear function, assembling perceptrons to create neural networks of multiple layers would not make sense anymore.

@@

\subsection{Activation functions}
\label{activation_functions}
\setlength{\marginparwidth}{2cm}\leavevmode \marginnote{\textbf{Johan}}
Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer. Activation functions must be non-linear in order to approximate extremely complex functions. In fact, neural networks are considered as universal approximators. Hornik et al. claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense" \cite{17}. According to Thomas Epelbaum \cite{18}, the most commonly used activation functions are:\\\\
 \noindent \textbf{Sigmoid function}\\
 The sigmoid function is defined as:
 \begin{equation}
  g(x) = \frac{1}{1+e^{-x}}
 \end{equation}
 Its derivative is:
 \begin{equation}
 g'(x) = g(x)(1-g(x))
 \end{equation}

@@

There exist multiple cost functions. For a regression problem, a commonly used loss function is the mean square error:
\begin{equation}
J(\hat{y}, y) = \frac{1}{m}[\sum_{i=1}^{m} (\hat{y}^{(i)} - y^{(i)})^{2}]
\end{equation}
For classification problems, the cross entropy function is regularly used. We distinguish the binary classification where the number of classes $n_{y}$ = 2 from the multiclass classification where $n_{y}$ > 2. In the case of binary classification, the cross entropy is:
\begin{equation}
J(\hat{y}, y) = -\frac{1}{m}\sum_{i=1}^{m} [y_{i}*log(\hat{y}_{i}) + (1-y_{i})*log(1-\hat{y_i})]
\end{equation}
In the case of multiclass classification, the categorical crossentropy is:
\begin{equation}
J(\hat{y}, y) = - \sum_{i=1}^{n_{y}} \sum_{j=1}^{m} (y_{ij}*log(\hat{y}_{ij}))
\end{equation}
Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weights in order to approach the minimum of the function. Therefore, deep learning problems can be considered as optimization problems. Solutions to these problems can be found using the gradient descent algorithm during backpropagation.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$. Generally speaking, "the intuition behind the backpropagation algorithm is as follows. Given a training example $(x^{(i)}, y^{(i)})$, we will first run a forward pass to compute all the activations throughout the network, including the output value of the network. Then, for each node $i$ in layer $l$, we would like to compute an "error term" $\partial^{(l)}_{i}$ that measures how much that node was "responsible" for any errors in our output. For an output node, we can directly measure the difference between the network’s activation and the true target value, and use that to define $\partial^{(n_{l})}_{i}$(where layer $n_{l}$ is the output layer). How about hidden units? For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.\\
In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting parameters of the model. The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network. The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ at point $x$ is given by:
\begin{equation}
\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{1}}]
\end{equation}
The gradient shows how much all parameters that constitute x need to change to minimize the function. In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$. The computation of all these gradients relies on the "chain rule". In the case of weights, the chain rule is:
\begin{equation}
\frac{\partial J}{\partial w_{jk}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial w_{jk}^{l}}
\end{equation}
Similarly, the chain rules has to be applied to the biases:
\begin{equation}
\frac{\partial J}{\partial b_{j}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial b_{j}^{l}}
\end{equation}
Once the gradients of each parameter are computed, these parameters are updated. The weights update is described by the following equation:
\begin{equation}
W^{[l]} = W^{[l]} - \alpha * \frac{\partial J}{\partial W^{[l]}}
\end{equation}
The biases update corresponds to:
\begin{equation}
b^{[l]} = b^{[l]} - \alpha * \frac{\partial J}{\partial b^{[l]}}
\end{equation}
%The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to manually be tuned.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Convolutional Neural Networks (CNN) are a specific type of deep neural networks. They are particular in that they contain layers which perform a mathematical operation named convolution on the input data. CNNs are mostly used in image and video analysis.\\
To perform a convolution, numerical input data and a "filter" are required. A filter can be seen as a $f$x$f$ numerical patch that moves across the entire input. It first moves horizontally until reaching the right-most border of the image. Then, it goes down a cell and starts from the border on the left-hand side. This process is repeated until the filter reaches the bottom-right corner of the image, which marks the end of the convolution. At each step, the dot product between the filter and the part of the input covered by the filter is computed. Figure \ref{fig:convolution} showcases a simple convolution which aims at finding vertical lines in a black and white image. For example, the output of the first step of the convolution (in red) is computed by evaluating the dot product between the blue filter and the red part of the input image: 
\begin{equation}
\begin{gathered}
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}0
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}= -2
\end{gathered}
\end{equation}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/convolution.png}
\caption{Convolutions - Basic convolution in a CNN}
\label{fig:convolution}
\end{figure}
\noindent Different parameters can change the way a convolution behaves. First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom. In this case, the so-called "stride" is equal to 1. Other applications could rely on a bigger stride. Furthermore, the previous example reduced the output size of $3$x$3$ in proportion to the initial input size of $5$x$5$. To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s. Three ways of padding images are commonly used as shown on figure \ref{fig:convolution_padding}:
\begin{itemize}
	\item \textbf{Valid:} The input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.

### 2020-03-05 Common/chapters/introduction.tex
Johan and Julien: Thesis layout updated until transfer learning :)

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}. An early detection and treatment increases the chances of recovery. In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses. Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.

Deep learning applications require a lot of data to perform well. While certain fields profit from massive amounts of publicly available data, the medical field is quite the opposite. First of all, medical information is protected by the doctor-patient confidentiality and cannot be shared freely. As a consequence, data first has to be collected, organized and anonymized. Then, additional information regarding the clinical significance of the samples, the location of the lesions, etc. must be provided so that deep learning models can make use of the data. This whole process is time consuming and medical institutions don't always see the benefits which could ensue from publishing datasets of good quality. As a consequence, the lack of data is one of the toughest challenge related to this field. In order to overcome it, techniques such as transfer learning exist. The latter aims at using independent but similar datasets in order to increase the performance of a model on a target dataset. In other words, models are trained on the former in order to learn relevant features which improve the results on the latter. In the case of cancer detection and classification, only few datasets are available for each body part. Therefore, the idea behind this work is to make use of datasets of different body parts to improve the classification of prostate lesions.

This work is divided into various parts. First, the related literature was reviewed in order to gather techniques and architectures which gave decent results.

Second, deep learning is presented from the ground up under the historical and technical points of view. This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible without any background in the field.

The next chapter deals with medical knowledge related to cancer. Characteristics of the disease are presented, before focusing on the various file formats which medical imaging data are exported into. The last part is critical since a lot of data processing was performed in order to be able to make use of these files.

Then, the architecture presented in a prostate cancer classification research paper was reproduced. This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.

Finally, the last part is dedicated to transfer learning and its application to cancer classification.

>>>>

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}. An early detection and treatment increases the chances of recovery. In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses. Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.\\
Deep learning applications require a lot of data to perform well. While certain fields profit from massive amounts of publicly available data, the medical field is quite the opposite. First of all, medical information is protected by the doctor-patient confidentiality and cannot be shared freely. As a consequence, data first has to be collected, organized and anonymized. Then, additional information regarding the clinical significance of the samples, the location of the lesions, etc. must be provided so that deep learning models can make use of the data. This whole process is time consuming and medical institutions don't always see the benefits which could ensue from publishing datasets of good quality. As a consequence, the lack of data is one of the toughest challenge related to this field. In order to overcome it, techniques such as transfer learning exist. The latter aims at using independent but similar datasets in order to increase the performance of a model on a target dataset. In other words, models are trained on the former in order to learn relevant features which improve the results on the latter. In the case of cancer detection and classification, only few datasets are available for each body part. Therefore, the idea behind this work is to make use of datasets of different body parts to improve the classification of prostate lesions.\\
This work is divided into various parts. First, the related literature was reviewed in order to gather techniques and architectures which gave decent results.\\
Second, deep learning is presented from the ground up under the historical and technical points of view. This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible without any background in the field.\\
The next chapter deals with medical knowledge related to cancer. Characteristics of the disease are presented, before focusing on the various file formats which medical imaging data are exported into. The last part is critical since a lot of data processing was performed in order to be able to make use of these files.\\
Then, the architecture presented in a prostate cancer classification research paper was reproduced. This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.\\
Finally, the last part is dedicated to transfer learning and its application to cancer classification.
